%! TEX root = ./main.tex
\documentclass[12pt]{article}

%--------Packages-------------
\usepackage{kyrem1sty}
%----------------------------


%--------Bibliography---------
\usepackage[backend=biber,style=alphabetic,doi=false,isbn=false,url=false,eprint=false]{biblatex}
\addbibresource{/home/kyrem1/Mathematics/bibs/comborefs.bib}
%----------------------------


%--------Hyper Setup-------
\hypersetup{%
  colorlinks=true,%
  linkcolor=blue,%
  citecolor=blue,%
  filecolor=blue,%
  menucolor=blue,%
  urlcolor=blue,%
  pdfnewwindow=true,%
  pdfstartview=FitBH
}   
%----------------------------


%--------Other Setup---------
\usepackage{ytableau} % For young diagrams
\usepackage{caption}
% Adjust caption spacing and font
\captionsetup{
  font=small, % Adjust the font size
  labelfont=bf,
  format=plain, % Use plain format to avoid any unwanted effects
  justification=raggedright, % Ensure the caption is justified, which can help with spacing
  singlelinecheck=false, % Applies justification setting even when the caption is a single line
}
%----------------------------


%--------Subfiles Setup-------
%\usepackage{subfiles}
%----------------------------


%--------Page Setup-----------
%\usepackage{geometry}\geometry{margin=1in}
\pagestyle{empty}%

\setlength{\hoffset}{-1.54cm}
\setlength{\voffset}{-1.54cm}

\setlength{\topmargin}{0pt}
\setlength{\headsep}{0pt}
\setlength{\headheight}{0pt}

\setlength{\oddsidemargin}{0pt}

\setlength{\textwidth}{195mm}
\setlength{\textheight}{250mm}
%----------------------------


%--------Metadata------------
\title{Intro Math Research Hw3}
\author{James Harbour}
%----------------------------


%--------Content-------------
\begin{document}

\maketitle


\section{Reading Comments}

\subsection*{QUOMODOCUMQUE Article}
\begin{itemize}
  \item I quite like the point about avoiding undermining your own work. Most advice I have heard for math talks have been about humility and being very specific about how much attention you put on your own work, but this warns against swinging the pendulum too far in the other direction.
  \item I entirely agree with the statement about timing; I pretty much unconsciously stop listening to a talk the moment the speaker goes over time.
\end{itemize}

\subsection*{Tao Article}
\begin{itemize}
  \item Tao's blogposts often have really poignant quotes, this one is no exception.
  \item I like that he expresses the importance of always giving a sense of global locality within a talk/paper, as this is often the main challenge for people who can already locally verify details. 
  \item It's very interesting the point that one should sit on a paper for a while before releasing to, say, the arXiv. 
\end{itemize}

\subsection*{Vakil Article}
  \textbf{Note:} I very much already admire Vakil as an expositor for his \textit{Rising Sea}. It's where I first learned category theory ages ago and I continue to reference it whenever I need to think about schemes.
\begin{itemize}
  \item When I initially began attending seminars, I found myself taking notes in a more ``traditional sense'' as Vakil puts it. I definitely agree with him now that this sapped from my focus on the talk and nowadays I take much less notes and only write key ideas.
  \item I find it quite interesting the idea of, after having written three things to take away down, cutting one of them to make room for another instead of simply adding to the list. This is quite cool as it forces you to distill the essence of the talk.
\end{itemize}

\newpage

\tableofcontents


\section{Project Topics}
\textbf{Assignment}: Look over the Project Topics document and start searching for readable
resources. List the two topics which look most interesting to you and
give two references for each of these topics (different from the one I gave).
Keep in mind the remark of ยง 2.2

\subsection*{Topic 1: Schur-Weyl Duality and its Analogues}
\begin{itemize}
  \item In \cite[Ch.4,6]{fultonharris}, Fulton and Harris provide a gentle introduction to Schur-Weyl duality and the various identities which arise as a consequence.
  \item In \cite{dipper:05}, Dipper, Doty, and Hu obtain an analogue of Schur-Weyl duality for symplectic groups using the Brauer algebra. 

\end{itemize}



\subsection*{Topic 2: Quasisymmetric Functions}
\begin{itemize}
  \item The book \textit{An Introduction to Quasisymmetric {S}chur Functions} \cite{qsymschur} by Luoto, Mykytiuk and van Willigenburg gives a broad overview of the machinery behind quasisymmetric functions (namely Hopf algebras) as well as a description of the basis of quasisymmetric {S}chur functions for the quasisymmetric function algebra.
  \item In \cite{stanley:95}, Stanley contstructs a symmetric function from a finite graph. This construction encodes the coloring data in a way that generalizes the graph chromatic polynomial. Nearly two decades later, Shareshian and Wachs in \cite{wachs:12} develop a quasisymmetric version of Stanley's chromatic function and establish connections to cohomological invariants coming from algebraic geometry. Three years later, Athanasiadis produced a wonderful elementary account (\cite{ath}) of the development of both the symmetric and quasisymmetric chromatic function . 
\end{itemize}



\newpage

\section{Symmetric Polynomials/Functions Exposition}

%\textbf{Assignment}: Revise your write-up on symmetric polynomials, focusing on the extra tip
%,in ยง 2.1 (``indispensable, interesting, illustrative.'') If you did not write about the basis of homogeneous symmetric
%functions last time, this should be included in your two pages. Add a third
%page dedicated to the combinatorial definition of Schur functions and their
%symmetry.

\subsection*{Preliminary Considerations} 
Throughout this article, fix a (unital) commutative ring $ R $ and a field $ k $. For simplicity, we work over vector spaces instead of general modules.

\begin{notation*}
  Let $ X $ be a set such that $ X = \{x_{i}\}_{i\in I} $ for some indexing set $ I $. By $ k[X] $ and $ k\llbracket X \rrbracket $, we denote the rings of (commutative) polynomials and power series (resp.) in indeterminates $ \{x_{i}\} $. We utilize multi-index notation throughout. Hence for $ \alpha_{\bigcdot}:I\to \N\cup\{0\} $ finitely supported, we write $ x_{\alpha} = \Pi_{i\in I}x_{i}^{\alpha_{i}} $ (where $ x_{i}^0 := 1 $ formally).
\end{notation*}

\subsection{Algebraic Background}

%Many common algebraic objects possess both a vector space structure and an internal product structure. For example, the set of $ n\times n $ matrices over $ k $, denoted $ M_{n}(k) $, has a product given by matrix multiplication and is also a vector space under addition and scalar multiplication by $ k $. 
%
%\begin{definition}\label{def:kalg}
%  Let $ A $ be a $ k $-vector space equipped with a map $ \cdot: A\times A\to A $ (written $ (x,y)\mapsto x\cdot y $). The pair $ (A,\cdot) $ is a \textit{$ k $-algebra} if, for $ x,y,z\in A $ and $ a,b\in k $, the following hold:
%  \begin{itemize}
%    \item $(x+y)\cdot z = x\cdot z + y\cdot z$,
%    \item $z\cdot (x+y) = z\cdot x + z\cdot y$,
%    \item $(a x) \cdot (b y) = (a b)(x\cdot y)$.
%  \end{itemize}
%\end{definition}
%%We abuse notation and often write $ xy $ for the product $ x\cdot y $. This is standard and usually clear from context. 
%%Observe that we did not assume the binary operation $ \cdot $ is associative. We will mainly focus on associative algebras; however, some non-associative examples (for culture) include Lie algebras, the Octonions, and $ \R^{3} $ equipped with the vector cross product.
%
%\begin{keyexample*}\label{ex:poly}
%  For $ X= \{x_{i}\}_{i\in I} $, the rings $ k[X] $ and $ k\llbracket X \rrbracket $ form the prototypical example of $ k $-algebras.
%\end{keyexample*}

Often in algebra, elements of a given object may be decomposed into a sum of simpler elements which are, in a sense, ``homogenous.'' For example, any polynomial in $ n $-variable may be decomposed into a sum of simpler polynomials each of which are futher sums of monomials of the same total degree.  In this way, a polynomial is split into a sum of homogenous parts. This behavior is codified in the notion of \textit{grading}.

\begin{definition}\label{def:gralg}
A \textit{graded $ k $-algebra} is a $ k $-algebra $ A $ together with a direct sum decomposition
\[
  A = \bigoplus_{i=0}^{\infty} A_{i}
\]
with $ A_{0},A_{1},\ldots $ vector spaces such that $ A_{i}\cdot A_{j}\sub A_{i+j} $ for all $ i,j\in \N\cup\{0\} $. For fixed $ i $, elements of $ A_{i} $ are called \textit{homogenous}. The choice of such a direct sum decomposition is a \textit{grading} for $ A $.
\end{definition}

%TODO insert exposition

\begin{keyexample*}
  As before, for $ X = \{x_{i}\}_{i\in I} $, we may give the ring $ k[X] $ a canonical grading by declaring $ A_{0}:=k $ and
  \[
    A_{n} := \Span_{k}\{x_{\alpha}: \alpha \text{ multi-index such that } \sum_{i\in I} \alpha_{i} = n\}.
  \]
\end{keyexample*} 
The reader is cautioned that not every $ k $-algebra has a nontrivial grading. In fact, it can be shown that the ring of formal power series $ k\llbracket x \rrbracket $ does not have a nontrivial grading.
%\begin{claim}
%  The algebra $ k\llbracket x \rrbracket $ does not possess a nontrivial grading.
%\end{claim}
%\begin{proof}[Proof of Claim]
%  Suppose 
%\end{proof}



\subsection{Symmetric Polynomials}

\begin{definition}\label{def:sympoly}
  The permutation group $ S_{n} $ acts naturally on the polynomial ring $ k[x_{1},\ldots,x_{n}] $ by defining $ \sigma \cdot x_{i_{1}}^{\alpha_1}\cdots x_{i_l}^{\alpha_{l}} := x_{\sigma(i_{1})}^{\alpha_{1}}\cdots x_{\sigma(i_{l})}^{\alpha_{l}}$ and extending by linearity. The ring of \textit{symmetric polynomials} in $ n $ indeterminates is the fixed points of this action, namely $ k[x_{1},\ldots, x_{n}]^{S_{n}} $.
\end{definition}

\subsection{Partitions and Compositions}
\begin{definition}\label{def:comp}\ 
  \begin{itemize}
    \item A \textit{partition} of $ n\in \N $ is a finite sequence $ \alpha=(\alpha_{1},\ldots, \alpha_{l}) $ of weakly decreasing positive integers which sum to $ n $. We denote the set of partitions of $ n $ by $ \Par(n) $. We denote the statement $ [\lambda\in \Par(n)] $ by $ \lambda\vdash n $. Also, we write $ \Par:=\bigcup_{n\geq0}\Par(n) $.
    \item A \textit{weak composition} of $ n\in \N $ is a (finitely supported) sequence $\alpha=(\alpha_{i})_{i=1}^{\infty}\in (\N\cup\{0\})^{\N}$ such that $ \sum_{i} \alpha_{i} = n $. The length of a weak composition $ \alpha $ is given by
  \[
    l(\alpha) :=\max\{i\in \N: \alpha_{i}\neq 0\}.
  \]
  \end{itemize}

  
\end{definition}

\begin{example}
  For $ n=5 $, the sequences $ \alpha=(1, 0, 2, 2, 0, 0, \ldots) $ and $ \beta=(2,0,1,2,0,0,\ldots) $ are distinct  weak compositions but neither are valid partitions of $ 5 $ due to the presence of a $ 0 $ between positive entries. On the other hand, $ \lambda = (2,2,1) $ is a partition of $ 5 $.
\end{example}

\subsection{Symmetric Functions}

\begin{definition}[pg. 308 in \cite{stanley2}]\label{def:symfunc}
    The ring $ \Lambda_{k} $ of symmetric functions over a field $ k $ is the subring of all $ f\in k\llbracket x_{1}, x_{2}, \ldots\rrbracket $ such that
    \[
      f(x_{\sigma(1)}, x_{\sigma(2)},\ldots) = f(x_{1},x_{2},\ldots) \text{ for all } \sigma\in \Sym(\N).
    \]
\end{definition}
\begin{remark}
    For the algebraically-minded, there is a more natural construction of $ \Lambda_{k} $ by viewing the ring as the colimit of a certain directed system of injections of symmetric polynomial rings
    \[
      k[x_{1},\ldots,x_{n}]^{S_{n}} \xhookrightarrow{\phi_{n}} k[x_{1},\ldots,x_{n+1}]^{S_{n+1}}.
    \]
    The construction of these maps $ \phi_{n} $ is somewhat involved. This does justify the intuition that a symmetric function is simply taking a symmetric polynomial and adding more data, as any element of a direct limit of inclusions is faithfully represented by an element of one of the constituent objects.

\end{remark}
\begin{definition}
    A symmetric function $ f\in \Lambda_{k} $ is homogenous of degree $ n $ if 
    \[
      f(x) = \sum_{\alpha\text{ weak composition of }n}c_{\alpha} x^{\alpha},
    \]
    where the $ c_{\alpha} $ are elements of $ k $. The set of degree $ n $ homogenous symmetric functions is denoted $ \Lambda_{k}^{n} $. these subspaces give $ \Lambda_{k} $ the structure of a graded $ k $-algebra, namely:
\end{definition}

\begin{itemize}
  \item Each $ \Lambda_{k}^{n} $ is a $ k $-vector space,
  \item $ \Lambda_{k}^{i} \Lambda_{k}^{j}\sub \Lambda_{k}^{i+j} $,
  \item $ \Lambda_{k}=\bigoplus_{n=0}^{\infty} \Lambda_{k}^{n} $ as $ k $-vector spaces.
\end{itemize}
The first interesting basis of $ \Lambda_{k} $ is the \textit{monomial symmetric functions}. Given $ \lambda\vdash n $, define $ m_{\lambda}\in \Lambda_{k}^{n} $ by 
\[
  m_{\lambda}:= \sum_{\alpha} x^{\alpha}
\]
where the sum is over all distinct permutations of the entries of $ \lambda $. The set $ \{m_{\lambda}: \lambda\vdash n\} $ forms a basis for $ \Lambda_{k}^{n} $, whence $\bigcup_{n\geq0} \{m_{\lambda}: \lambda\vdash n\} = \{m_{\lambda}: \lambda\in\Par\} $ forms a basis for $ \Lambda_{k} $.

\subsection{Complete Homogenous Symmetric Functions}

From the monomial symmetric functions, we may form another interesting basis for $ \Lambda_{k} $ called the \textit{complete homogenous symmetric functions} $ h_{\lambda} $ by setting 
\[
  h_{\lambda}:=\prod_{i=1}^{\infty} \sum_{\nu\vdash \lambda_{i}} m_{\nu}.
\]
where $ \lambda = (\lambda_{1},\lambda_{2},\ldots) $. Again, the set $ \{h_{\lambda}: \lambda\vdash n\} $ is a basis for $ \Lambda_{k}^{n} $ and the set $ \{h_{\lambda}: \lambda\in\Par\} $ is a basis for $ \Lambda_{k} $.
\subsection{Schur Functions}

\begin{definition}\label{def:youngdiagram}
  Given $ \lambda\vdash n $, the \textit{Ferrers diagram of shape $ \lambda $} is the set $\{(i,j)\in \N^{2} : j\in \N, 1\leq i \leq \lambda_{j}\}$ depicted as points in $ \R^{2} $. The \textit{Young diagram of shape $ \lambda $} is depicted identically to the Ferrers diagram except the points are replaced with squares. The \textit{size} of the diagram is the number of entries, namely $ n $. We depict the case $ (5,2,1)\vdash 8 $ below.
\end{definition}

\begin{figure}[h]
\centering
% Align tableaux at the top of their respective minipages
\ytableausetup{boxframe=0pt, aligntableaux=top}
\begin{minipage}[t]{0.20\textwidth}
\centering
\ydiagram[\bullet]{1,2,5}
% Adjust space below to align captions uniformly
\vspace{5pt} % Adjust the space to ensure alignment if necessary
\caption*{\textit{Ferrers Diagram} for the partition (5,2,1)}
\end{minipage}
\quad % This adds some space between the diagrams
\begin{minipage}[t]{0.20\textwidth}
\centering
\ytableausetup{boxframe=normal, aligntableaux=top}
\begin{ytableau}
*(white)\\
*(white) & *(white) \\
*(white) & *(white) & *(white) & *(white) & *(white) \\
\end{ytableau}
\vspace{5pt} % Keep consistent vertical space
\caption*{\textit{Young Diagram} for the partition (5,2,1)}
\end{minipage}
\quad
\begin{minipage}[t]{0.20\textwidth}
\centering
\begin{ytableau}
9 \\
5 & 5 \\
1 & 1 & 1 & 4 & 6 \\
\end{ytableau}
\vspace{5pt} % Adjust as needed
\caption*{\textit{Semi-standard Young tableau} for the partition (5,2,1)}
\end{minipage} 
\quad
\begin{minipage}[t]{0.20\textwidth}
\centering
\begin{ytableau}
8 \\
5 & 7 \\
1 & 2 & 3 & 4 & 6 \\
\end{ytableau}
\vspace{5pt} % Uniform spacing for alignment
\caption*{\textit{Standard Young tableau} for the partition (5,2,1)}
\end{minipage} 
\end{figure}


\begin{definition}\label{def:youngtableaux}
  Given $ \lambda\vdash n $ and a Young diagram of shape $ \lambda $, a \textit{semi-standard Young tableau of shape $ \lambda $} is a filling of the boxes of the Young diagram with positive integers such that 
  \begin{itemize}
    \item the entries are weakly increasing along rows,
    \item the entries are strictly increasing up columns.
  \end{itemize}
  A semi-standard Young tableau of size $ n $ is said to be \textit{standard} if the elements of $ \{1,\ldots,n\} $ each appear exactly once in the tableau. We write $ SSYT(\lambda) $ and $ SYT(\lambda) $ for the sets of semi-standard and standard Young tableaux of shape $ \lambda $. Given a semi-standard Young tableaux $ \mathcal{T} $, the \textit{weight} of $ \mathcal{T} $ is a function $ \alpha=\alpha_{\mathcal{T}}:\N \to \N$ given by
  \[
    \alpha(i) := \text{number of times $ i $ appears in $ \mathcal{T} $}.
  \]
  Note that $ \alpha(i) = 0 $ for sufficiently large $ i $, so we may write $ x^{\alpha} = x_{1}^{\alpha(1)}x_{2}^{\alpha(2)} \cdots $ and obtain a valid monomial.
\end{definition}


\begin{definition}\label{def:schur}
  Given $ \lambda\vdash n $, the \textit{Schur function} indexed by $ \lambda $ is defined by
 \[
  s_{\lambda} := \sum_{\mathcal{T}\in SSYT(\lambda) } x^{\alpha_{\mathcal{T}}}
\] 
where $ \alpha_{\mathcal{T}} $ denotes the weight of the Young tableau $ \alpha $.
\end{definition}




\end{document}
