%! TEX root = ./main.tex
\documentclass[12pt]{article}

%--------Packages-------------
\usepackage{kyrem1sty}
%----------------------------


%--------Bibliography---------
\usepackage[backend=biber,style=alphabetic,doi=false,isbn=false,url=false,eprint=false]{biblatex}
\addbibresource{/home/kyrem1/Mathematics/bibs/comborefs.bib}
%----------------------------


%--------Hyper Setup-------
\hypersetup{%
  colorlinks=true,%
  linkcolor=blue,%
  citecolor=blue,%
  filecolor=blue,%
  menucolor=blue,%
  urlcolor=blue,%
  pdfnewwindow=true,%
  pdfstartview=FitBH
}   
%----------------------------


%--------Subfiles Setup-------
%\usepackage{subfiles}
%----------------------------


%--------Page Setup-----------
%\usepackage{geometry}\geometry{margin=1in}
\pagestyle{empty}%

\setlength{\hoffset}{-1.54cm}
\setlength{\voffset}{-1.54cm}

\setlength{\topmargin}{0pt}
\setlength{\headsep}{0pt}
\setlength{\headheight}{0pt}

\setlength{\oddsidemargin}{0pt}

\setlength{\textwidth}{195mm}
\setlength{\textheight}{250mm}
%----------------------------


%--------Metadata------------
\title{Intro Math Research Hw2}
\author{James Harbour}
%----------------------------


%--------Content-------------
\begin{document}

\maketitle
\tableofcontents


\section{Reading Comments}

\subsection*{Pak Reading}


\subsection*{Godsil Reading}


\newpage

\section{Symmetric Polynomials/Functions Exposition}

%\textbf{Assignment}: Write two pages about symmetric polynomials, confining yourself to what
%we learned in class (definition of the algebra, the monomial basis, and partitions). The only new material should be a discussion on the homogenous
%symmetric functions.

\subsection*{Preliminary Considerations} 
Throughout this article, fix a (unital) commutative ring $ R $ and a field $ k $. For simplicity, we work over vector spaces instead of general modules.

\begin{notation*}
  Let $ X $ be a set such that $ X = \{x_{i}\}_{i\in I} $ for some indexing set $ I $. By $ k[X] $ and $ k\llbracket X \rrbracket $, we denote the rings of (commutative) polynomials and power series (resp.) in indeterminates $ \{x_{i}\} $. We utilize multi-index notation throughout. Hence for $ \alpha_{\bigcdot}:I\to \N\cup\{0\} $ finitely supported, we write $ x_{\alpha} = \Pi_{i\in I}x_{i}^{\alpha_{i}} $ (where $ x_{i}^0 := 1 $ formally).
\end{notation*}

\subsection{Algebraic Background}

Many common algebraic objects possess both a vector space structure and an internal product structure. For example, the set of $ n\times n $ matrices over $ k $, denoted $ M_{n}(k) $, has a product given by matrix multiplication and is also a vector space under addition and scalar multiplication by $ k $. 

\begin{definition}\label{def:kalg}
  Let $ A $ be a $ k $-vector space equipped with a map $ \cdot: A\times A\to A $ (written $ (x,y)\mapsto x\cdot y $). The pair $ (A,\cdot) $ is a \textit{$ k $-algebra} if, for $ x,y,z\in A $ and $ a,b\in k $, the following hold:
  \begin{itemize}
    \item $(x+y)\cdot z = x\cdot z + y\cdot z$,
    \item $z\cdot (x+y) = z\cdot x + z\cdot y$,
    \item $(a x) \cdot (b y) = (a b)(x\cdot y)$.
  \end{itemize}
\end{definition}
%We abuse notation and often write $ xy $ for the product $ x\cdot y $. This is standard and usually clear from context. 
%Observe that we did not assume the binary operation $ \cdot $ is associative. We will mainly focus on associative algebras; however, some non-associative examples (for culture) include Lie algebras, the Octonions, and $ \R^{3} $ equipped with the vector cross product.

\begin{keyexample*}\label{ex:poly}
  For $ X= \{x_{i}\}_{i\in I} $, the rings $ k[X] $ and $ k\llbracket X \rrbracket $ form the prototypical example of $ k $-algebras.
\end{keyexample*}

Often in algebra, elements of a given object may be decomposed into a sum of simpler elements which are, in a sense, ``homogenous.'' For example, any polynomial in $ n $-variable may be decomposed into a sum of simpler polynomials each of which are futher sums of monomials of the same total degree.  In this way, a polynomial is split into a sum of homogenous parts. This behavior is codified in the notion of \textit{grading}.

\begin{definition}\label{def:gralg}
A \textit{graded $ k $-algebra} is a $ k $-algebra $ A $ together with a direct sum decomposition
\[
  A = \bigoplus_{i=0}^{\infty} A_{i}
\]
with $ A_{0},A_{1},\ldots $ vector spaces such that $ A_{i}\cdot A_{j}\sub A_{i+j} $ for all $ i,j\in \N\cup\{0\} $. For fixed $ i $, elements of $ A_{i} $ are called \textit{homogenous}. The choice of such a direct sum decomposition is a \textit{grading} for $ A $.
\end{definition}

%TODO insert exposition

\begin{keyexample*}
  As before, for $ X = \{x_{i}\}_{i\in I} $, we may give the ring $ k[X] $ a canonical grading by declaring $ A_{0}:=k $ and
  \[
    A_{n} := \Span_{k}\{x_{\alpha}: \alpha \text{ multi-index such that } \sum_{i\in I} \alpha_{i} = n\}.
  \]
\end{keyexample*} 
The reader is cautioned that not every $ k $-algebra has a nontrivial grading. In fact, it can be shown that the ring of formal power series $ k\llbracket x \rrbracket $ does not have a nontrivial grading.
%\begin{claim}
%  The algebra $ k\llbracket x \rrbracket $ does not possess a nontrivial grading.
%\end{claim}
%\begin{proof}[Proof of Claim]
%  Suppose 
%\end{proof}



\subsection{Symmetric Polynomials}

\begin{definition}\label{def:sympoly}
  The permutation group $ S_{n} $ acts naturally on the polynomial ring $ k[x_{1},\ldots,x_{n}] $ by defining $ \sigma \cdot x_{i_{1}}^{\alpha_1}\cdots x_{i_l}^{\alpha_{l}} := x_{\sigma(i_{1})}^{\alpha_{1}}\cdots x_{\sigma(i_{l})}^{\alpha_{l}}$ and extending by linearity. The ring of \textit{symmetric polynomials} in $ n $ indeterminates is the fixed points of this action, namely $ k[x_{1},\ldots, x_{n}]^{S_{n}} $.
\end{definition}

\subsection{Compositions and Partitions}
\begin{definition}\label{def:comp}\ 
  \begin{itemize}
    \item A \textit{partition} of $ n\in \N $ is a set $ \alpha=\{\alpha_{1},\ldots, \alpha_{l}\} $ of positive integers which sum to $ n $. We denote the set of partitions of $ n $ by $ \Par(n) $. We denote the statement $ [\lambda\in \Par(n)] $ by $ \lambda\vdash n $. Also, we write $ \Par:=\bigcup_{n\geq0}\Par(n) $.
    \item A \textit{weak composition} of $ n\in \N $ is a (finitely supported) sequence $\alpha=(\alpha_{i})_{i=1}^{\infty}\in (\N\cup\{0\})^{\N}$ such that $ \sum_{i} \alpha_{i} = n $. The length of a weak composition $ \alpha $ is given by
  \[
    l(\alpha) :=\max\{i\in \N: \alpha_{i}\neq 0\}.
  \]
  \end{itemize}

  
\end{definition}

\subsection{Symmetric Functions}

\begin{definition}[pg. 308 in \cite{stanley2}]\label{def:symfunc}
    The ring $ \Lambda_{k} $ of symmetric functions over a field $ k $ is the subring of all $ f\in k\llbracket x_{1}, x_{2}, \ldots\rrbracket $ such that
    \[
      f(x_{\sigma(1)}, x_{\sigma(2)},\ldots) = f(x_{1},x_{2},\ldots) \text{ for all } \sigma\in \Sym(\N).
    \]
\end{definition}
\begin{remark}
    For the algebraically-minded, there is a more natural construction of $ \Lambda_{k} $ by viewing the ring as the colimit of a certain directed system of inclusions of symmetric polynomial rings
    \[
      k[x_{1},\ldots,x_{n}]^{S_{n}} \xhookrightarrow{\phi_{n}} k[x_{1},\ldots,x_{n+1}]^{S_{n+1}}.
    \]
    The construction of these inclusions $ \phi_{n} $ is somewhat involved. This does justify the intuition that a symmetric function is simply taking a symmetric polynomial and adding more data, as any element of a direct limit of inclusions is faithfully represented by an element of one of the constituent objects.

\end{remark}
\begin{definition}
    A symmetric function $ f\in \Lambda_{k} $ is homogenous of degree $ n $ if 
    \[
      f(x) = \sum_{\alpha\text{ weak composition of }n}c_{\alpha} x^{\alpha},
    \]
    where the $ c_{\alpha} $ are elements of $ k $. The set of degree $ n $ homogenous symmetric functions is denoted $ \Lambda_{k}^{n} $. these subspaces give $ \Lambda_{k} $ the structure of a graded $ k $-algebra, namely:
\end{definition}

\begin{itemize}
  \item Each $ \Lambda_{k}^{n} $ is a $ k $-vector space,
  \item $ \Lambda_{k}^{i} \Lambda_{k}^{j}\sub \Lambda_{k}^{i+j} $,
  \item $ \Lambda_{k}=\bigoplus_{n=0}^{\infty} \Lambda_{k}^{n} $ as $ k $-vector spaces.
\end{itemize}
The first interesting basis of $ \Lambda_{k} $ is the \textit{monomial symmetric functions}. Given $ \lambda\vdash n $, define $ m_{\lambda}\in \Lambda_{k}^{n} $ by 
\[
  m_{\lambda}:= \sum_{\alpha} x^{\alpha}
\]
where the sum is over all distinct permutations of the entries of $ \lambda $. The set $ \{m_{\lambda}: \lambda\vdash n\} $ forms a basis for $ \Lambda_{k}^{n} $, whence $\bigcup_{n\geq0} \{m_{\lambda}: \lambda\vdash n\} = \{m_{\lambda}: \lambda\in\Par\} $ forms a basis for $ \Lambda_{k} $.

\subsection{Complete Homogenous Symmetric Functions}

From the monomial symmetric functions, we may form another interesting basis for $ \Lambda_{k} $ called the \textit{complete homogenous symmetric functions} $ h_{\lambda} $ by setting 
\[
  h_{\lambda}:=\prod_{i=1}^{\infty} \sum_{\nu\vdash \lambda_{i}} m_{\nu}.
\]
where $ \lambda = (\lambda_{1},\lambda_{2},\ldots) $. Again, the set $ \{h_{\lambda}: \lambda\vdash n\} $ is a basis for $ \Lambda_{k}^{n} $ and the set $ \{h_{\lambda}: \lambda\in\Par\} $ is a basis for $ \Lambda_{k} $.
\end{document}
